---
title: "Análisis de la TRN de C.salexigens"
author: "Daniel Ruiz Palacios"
date: ""
output:
  html_document:
    toc: yes
    toc_depth: 2
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: 2
location: Universidad de Sevilla
---
# Steps prior to the analysis:

## 1) In Cytoscape, open GRN saved as .csv as following: File --> Import --> Network from File...

## 2) Save network as .sif as following: Export as Network --> .sif

## 3) Save networks at the same directory that this script:

## 4) Load networks:

```{r}

library(this.path)
library(igraph)
library(gsubfn)

# Get current directory
cur_dir2 = dirname(this.path())
setwd(cur_dir2)
getwd()

# Make an object with the name of all GRN
files <- list.files(path=cur_dir2, pattern="*.sif", full.names=TRUE, recursive=FALSE)

data <- data.frame(gsub(".*/","",files))
data <- gsub(".*/","",files)
data <- data.frame(gsub("\\..*","",data))

colnames(data) <- c('GRN_name')
```
# GRNs analysis

## 1) Number of nodes and interactions of each network.
```{r}

nodes <- list()
interactions <- list()

for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  nodes[[i]] <- gorder(graph = assign(i,network))
  interactions[[i]] <- gsize(graph = assign(i,network))
}

names(nodes) <- data$GRN_name
names(interactions) <- data$GRN_name

```

## 2) Analysis of global parameters of GRNs: 

### 2.1) Diameter:

```{r}

diam <- list()

# Maximum of all possible distances between the nodes constituting the GRN
for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  diam[[i]] <- diameter(graph = assign(i,network), unconnected = TRUE)
}

names(diam) <- data$GRN_name

```

### 2.2) Radius:

```{r}

rad <- list()

# Eccentricity: Maximum distances between a node and the rest of nodes in the network

# Radius: Minimum of all the eccentricities of the nodes that constitute a network

for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  rad[[i]] <- radius(graph = assign(i,network), mode="all")
}

names(rad) <- data$GRN_name

```

### 2.3) Distribution of nodes degree:

```{r}

node.degree.min <- list()
node.degree.max <- list()
for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  
  # Minimum degree of all nodes
  node.degree.min[[i]] <- min(degree(graph = assign(i,network), loops = F))
  # Maximum degree of all nodes
  node.degree.max[[i]] <- max(degree(graph = assign(i,network), loops = F))
  # Graphical representation of nodes degree
  hist(degree(graph = assign(i,network), loops = F), freq = F, xlim = range(0,1000))
}

names(node.degree.min) <- data$GRN_name
names(node.degree.max) <- data$GRN_name

# As we can see, the graph of nodes degree follows a negative potential graph.

```

### 2.4) Average path length

```{r}
apl <- list()
for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  apl[[i]] <- average.path.length(graph = assign(i,network))
}

names(apl) <- data$GRN_name

```

### 2.5) Transitivity or clustering coefficient:

```{r}

# Transitivity measures the probability that the adjacent vertices of a vertex are connected. This is sometimes also called a grouping coefficient.

av.tran <- list()

for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  av.tran[[i]] <- transitivity(graph = assign(i,network), type = 'average')
}

names(av.tran) <- data$GRN_name

```

## 3) Scale-free features:
### 3.1) Checking whether the network is scale-free using linear regression:

```{r}

adj.r.squared <- list()
p_value <- list()

for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  
  # Adj. R^2 
  
  adj.r.squared[[i]] <- summary(lm(log2(as.vector(table(degree(graph = assign(i,network))))) ~ log2(as.numeric(names(table(degree(graph = assign(i,network))))))))$adj.r.squared
  
  # p-value
  
  p_value[[i]] <- summary(lm(log2(as.vector(table(degree(graph = assign(i,network))))) ~ log2(as.numeric(names(table(degree(graph = assign(i,network))))))))$coefficients[2,4]
}

names(adj.r.squared) <- data$GRN_name
names(p_value) <- data$GRN_name


# R-squared es elevado y p-valor es bajo --> Es posible que sea una red libre de escala
```

R-squared is high and p-value is low --> It is a scale-free network. As the network is scale-free, it has a number of properties that are:

### 3.2) Hubs:

```{r}
# Hubs are nodes whose node degree is greater than 0.95

# 0.95 hubs
hubs_0.95 <- list()

for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  hubs_0.95[[i]] <- length(which(degree(assign(i,network)) > quantile(degree(assign(i,network)),probs = 0.95)))
}

# 0.99 hubs (More restrictive)

hubs_0.99 <- list()
for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  hubs_0.99[[i]] <- length(which(degree(assign(i,network)) > quantile(degree(assign(i,network)),probs = 0.99)))
}

names(hubs_0.95) <- data$GRN_name
names(hubs_0.99) <- data$GRN_name

```

### 3.3) All TFs and highest TF for each network:

```{r}
highest_TF <- list()
networks_TF <- list()

for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  hubs95 <- data.frame(which(degree(assign(i,network)) > quantile(degree(assign(i,network)),probs = 0.95)))
  
for (j in hubs95) {
  networks_TF[[i]] <- sort(degree(graph = network)[j], decreasing = T)
  highest_TF[[i]] <- names(sort(degree(graph = network)[j], decreasing = T)[1])
}
  
}

names(networks_TF) <- data$GRN_name
networks_TF <- unlist(networks_TF)

TF.names <- gsub(".*\\.","",names(networks_TF))
networks_names <- gsub("\\..*","",names(networks_TF))
nodes.degree <- unname(networks_TF)

networks_TF <- as.data.frame(TF.names)
networks_TF["node degree"] <- as.data.frame(nodes.degree)
networks_TF["networks_names"] <- as.data.frame(networks_names)

write.csv2(networks_TF, file = "all_networks_TF.txt")

names(highest_TF) <- data$GRN_name

```

### Checking if GRNs are small-world networks :

```{r}

# apl < ln(N) --> TRUE --> Small-world networks

library("SciViews")

logarithm_N <- list()
logarithm_logarithm_N <- list()

for (i in files){
  TRN <- read.csv(i, sep = '\t', header = F)
  TRN[,2] = NULL
  TRN <- unique(TRN)
  network <- graph.data.frame(assign(i,TRN),directed = F)
  logarithm_N[[i]] <- (ln(gorder(graph = assign(i,network))))
  logarithm_logarithm_N[[i]] <- (ln(ln(gorder(graph = assign(i,network)))))
}

names(logarithm_N) <- data$GRN_name
names(logarithm_logarithm_N) <- data$GRN_name

```

# All data together:

```{r}
data["Nodes"] <- t(as.data.frame(nodes))
data["Interactions"] <- t(as.data.frame(interactions))
data['Diameter'] <- t(as.data.frame(diam))
data['Radius'] <- t(as.data.frame(rad))
data['Minimum node degree'] <- t(as.data.frame(node.degree.min))
data["Maximum node degree"] <- t(as.data.frame(node.degree.max))
data["TF with MND"] <- t(as.data.frame(highest_TF))
data["R squared"] <- t(as.data.frame(adj.r.squared))
data["p-value"] <- t(as.data.frame(p_value))
data["Average path length"] <- t(as.data.frame(apl))
data["ln(N)"] <- t(as.data.frame(logarithm_N))
data["ln(ln(N))"] <- t(as.data.frame(logarithm_logarithm_N))
data["Transitivity"] <- t(as.data.frame(av.tran))
data["Hubs"] <- t(as.data.frame(hubs_0.95))

rownames(data) <- data$GRN_name
write.csv2(x = data, file = "13_networks_data.txt")

```

# Comparison of the characteristics of the networks obtained from Abasy Atlas and the networks of C.salexigens.

```{r}

data$`TF with MND` <- NULL

rownames(data) <- data$GRN_name
data.abasy.atlas <- data[c(5,6,7,8,9,10,11,12,13),]
data.old.network.only.orthology.c.sal <- data[1,]
data.old.network.c.sal <- data[2,]
data.new.network.c.sal <- data[3,]
data.final.network.c.sal <- data[4,]

data["TF with MND"] <- t(as.data.frame(highest_TF))
data <- data[,c(1,2,3,4,5,6,7,15,8,9,10,11,12,13,14)]

data.abasy.atlas$GRN_name <- NULL
data.old.network.c.sal$GRN_name <- NULL
data.old.network.only.orthology.c.sal$GRN_name <- NULL
data.new.network.c.sal$GRN_name <- NULL
data.final.network.c.sal$GRN_name <- NULL

data.mean <- as.data.frame(colMeans(data.abasy.atlas))
data.mean[,2] <- as.data.frame(colMeans(data.old.network.c.sal))
data.mean[,3] <- as.data.frame(colMeans(data.old.network.only.orthology.c.sal))
data.mean[,4] <- as.data.frame(colMeans(data.new.network.c.sal))
data.mean[,5] <- as.data.frame(colMeans(data.final.network.c.sal))
colnames(data.mean) <- c("mean_abasy_atlas","old_network_c_sal","orthologous_old_network","orthologous_new_network","final_network_c_sal")
data.mean <- t(data.mean)
data
data.mean

write.csv2(x = data.mean, file = "abasy_old_new_final_data.txt")
```